Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-09-05 Paper/method.tex
..

In this section, we present a generalized framework for computing the
active subspace. The framework is employed to the H$_2$/O$_2$ reaction kinetics
problem whereby the 
pre-exponent ($A_i$) in the rate law associated with individual reactions provided in
Table~\ref{tab:kinetics} is considered to be uniformly
distributed in the interval, $[0.9A_i^\ast, 1.1A_i^\ast]$; $A_i^\ast$ is the nominal
estimate provided in~\cite{Yetter:1991}.
Specifically, we employ two strategies: 
A gradient-based strategy that involves computation of model
gradients in order to construct the matrix $\hat{\mat{C}}$
in~\eqref{eq:chat}, and a gradient-free approach 
that involves a regression-based local linear approximation of
the model output to circumvent the computational cost associated with gradient
computation. 
The gradient-based approach is based on Algorithm
1.1, and the gradient-free approach is based on Algorithm 1.2
in~\cite{Constantine:2015}. The two strategies are implemented in an
iterative manner to ensure computational efficiency by avoiding 
extraneous model evaluations once a converged active subspace
is achieved.

@@

\bigskip
\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{An iterative gradient-based approach for discovering the active subspace}
  \begin{algorithmic}[1]
\Require $\theta_l$, $\theta_u$, $\tau$. 
\Ensure $\hat{\mat{\Lambda}}$, $\hat{\mat{W}}$, $\bm{\nu}_r(f)$ %$\eta$. %
    \Procedure{Gradient-based}{}
    \State Set $k$ = 0
	\State Draw $n_k$ random samples, $\{\bm{\xi}_i\}_{i=1}^{n_k}$ 
         according to $\pi_{\bm{\xi}}$. 
    \State Set $N_\text{total}$ = $n_k$ 
	\iffalse \State Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_r}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_r}$ \fi
	\State For each $i=1, \ldots, N_\text{total}$, compute $f(\bm{\xi}_i)$ and the gradient $\bm{g}^i = \nabla_{\bm{\xi}}f(\bm{\xi}_i)$
    \iffalse  
	\Statex\hspace{5mm} Using Finite Difference: 
	\Statex\hspace{5mm} i. Assign a small increment, $d\xi$.
	\Statex\hspace{5mm} ii. Augment the set of samples with neighboring points.
	\be \{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}:~\{\bm{\xi}_k\}_{k=1}^{n_1} \cup
        \{\xi_{k,j}+d\xi\}_{j=1}^{N_p} \nonumber
	\ee
	\Statex\hspace{5mm} iii. Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}$
	\Statex\hspace{5mm} iv. Using the augmented set, $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}$
         compute $\bm{g}^i$.
         \fi 
	\State Compute $\hat{\mat{C}}$ and its eigenvalue decomposition 
		$\hat{\mat{C}}$= $\frac{1}{N_\text{total}}\sum\limits_{i=1}^{N_\text{total}}[\bm{g}^i][\bm{g}^i]^\top$ = 
		$\hat{\mat{W}}^{(k)}\hat{\mat{\Lambda}}^{(k)} \hat{\mat{W}}^{(k)\top}$
	%\State Eigenvalue decomposition, $\mat{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$%
	\State Partition: $\hat{\mat{\Lambda}}^{(k)}=
        \begin{bmatrix} \hat{\mat{\Lambda}}_1^{(k)} & \\ & \hat{\mat{\Lambda}}_2^{(k)} \end{bmatrix}$, 
        $\hat{\mat{W}}^{(k)}=\begin{bmatrix} \hat{\mat{W}}_1^{(k)} & \hat{\mat{W}}_2^{(k)} \end{bmatrix}$, 
        $\hat{\mat{\Lambda}}_1^{(k)}\in \mathbb{R}^{N_p\times r}$
	\Loop
		\State Set $k$ = $k$ + 1
		\State Draw $n_k =  \lceil\beta n_{k-1}\rceil$  new random samples 
                $\{\bm{\xi}_i\}_{i=1}^{n_k}$  $\beta\in[0,1]$  $k = n_{k-1}+1,\ldots,n_{k-1}+n_k$.

	%	\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.%
		\State Set $N_\text{total}$ = $N_\text{total}$ + $n_k$ 
		\State Compute $\bm{g}^i = \nabla_{\bm{\xi}_i}f(\bm{\xi}_i)$, 
             	$i=n_{k-1}+1, \ldots, n_{k-1}+n_k$.  
		\State Compute $\hat{\mat{C}}$ = 
        	$\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^i][\bm{g}^i]^\top$
		\State Eigenvalue decomposition, $\hat{\mat{C}}$ = $\hat{\mat{W}}^{(k)}\hat{\mat{\Lambda}}^{(k)}
		 \hat{\mat{W}}^{(k)\top}$
		\State Compute $\delta \hat{\mat{W}}_{1,j}^{(k)}$ = 
                       \scalebox{1.25}{$\frac{\|(\hat{\mat{W}}_{1,j}^{k})^2 - 
                       (\hat{\mat{W}}_{1,j}^{k-1})^2\|}{\|(\hat{\mat{W}}_{1,j}^{k-1})^2\|}$}, 
                       $j = 1,\ldots,r$.
		\If {$\max\left(\delta \hat{\mat{W}}_{1,j}^{(k)}\right)<\tau$}
			\State break
		\EndIf
	\EndLoop
	\State Compute $\nu_{i,r}(f) = \sum\limits_{j=1}^{r} \Lambda_j w_{i,j}^2$,
	$i=1,\ldots,N_p$.
	\State Normalize $\nu_{i,r}(f)$ as $\tilde{\nu}_{i,r}(f)$ = \scalebox{1.25}{$\frac{\nu_{i,r}(f)}{\sum_i\nu_{i,r}(f)}$}.

@@

In the gradient-free approach, the model output is first computed at an initial set of $n_0$ samples. 
An independent set of $M$ samples is also drawn according to the joint probability distribution of the inputs.
For each sample in $M$, $p$ nearest neighbors in the $n_0$ set are identified. A least-squares fit to model
evaluations at the $p$ samples is performed. The slope vector ($\bm{b}_i$) of the regression-fit is stored
separately and the process is repeated for each sample in $M$. The symmetric and positive semi-definite matrix,
$\hat{\mat{C}}$ is constructed using the slope vectors as follows:
%
\be
\hat{\mat{C}} = \frac{1}{M}\sum\limits_{i=1}^{M}\bm{b}_i\bm{b}_i^\top = \hat{\bm{W}}\hat{\bm{\Lambda}}\hat{\bm{W}}^\top
\ee
%
Hence, the gradient-free approach essentially differs in the procedure for constructing the matrix, $\hat{\mat{C}}$.
Instead of the model evaluations, a linear regression-based approximation is used to estimate the gradient of
the model output. The suitability of this approach thus depends upon the accuracy of this approximation. 
The active subspace computation involves the same sequence of steps as outlined previously for the gradient-based 
approach in Algorithm~\ref{alg:grad}. The procedure for constructing $\hat{\mat{C}}$ using a regression-based
local linear approximation
of the model output is provided in Algorithm~\ref{alg:free2}.

%\bigskip
%\begin{breakablealgorithm}
%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
%\renewcommand{\algorithmicensure}{\textbf{Output:}}
%  \caption{An iterative gradient-free approach for discovering the active subspace}
%  \begin{algorithmic}[1]
%\Require $\theta_l$, $\theta_u$, $M$, $\tau$. 
%\Ensure $\Lambda$, $W$, $\alpha$ %$\eta$. 
%    \Procedure{Gradient-free}{}
%    \State Set $k$ = 0
%	\State Draw $n_k$ random samples, $\{\bm{\xi}_j\}_{j=1}^{n_k}$
%	\State Set $N_\text{total}$ = $n_k$ 
%	\iffalse \State Project to the physical space:
%        $\{\bm{\theta}_k\}_{i=1}^{n_k}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_r}$
%        \fi
%     \State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ 
%	\State Compute $f(\bm\xi_j)$, $j=1, \ldots, N_\text{total}$.
%	\State Compute the matrix $\mathbb{C}$ and its eigenvalue decomposition using Algorithm 3
%	%\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
%	\State Partition: $\Lambda^{(k)}~=~ 
%        \begin{bmatrix} \Lambda_1^{(k)} & \\ & \Lambda_2^{(k)} \end{bmatrix}$, 
%        $W^{(k)}~=~\begin{bmatrix} W_1^{(k)} & W_2^{(k)} \end{bmatrix}$, 
%        $\Lambda_1^{(k)}\in \mathbb{R}^{d\times\mathcal{S}}$
%	\Loop
%		\State Set $k$ = $k$ + 1
%		\State Draw $n_k =  \beta n_{k-1}$  new random samples 
%		$\{\bm{\xi}_j\}_{j=1}^{n_k}$  $\beta =??$  $j = n_{k-1}+1,\ldots,n_{k-1}+n_k$. 
%        \State $N_\text{total}$ = $N_\text{total}$ + $n_k$    
%	\iffalse	\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$. \fi
%		 
%		\State Compute $f(\bm\xi_j)$, $j=n_{k-1}+1, \ldots, n_{k-1}+n_k$.  
%		\State Construct the matrix, $\mat{C}$ and its eigenvalue decomposition
%		\State Compute $\delta W_{1,j}^{(k)}$ = 
%                       \scalebox{1.25}{$\frac{\|(W_{1,j}^{k})^2 - 
%                       (W_{1,j}^{k-1})^2\|}{\|(W_{1,j}^{k-1})^2\|}$}, 
%                       $j = 1,\ldots,\mathcal{S}$.
%		\If {$\max\left(\delta W_{1,j}^{(k)}\right)<\tau$}
%			\State break
%		\EndIf
%	\EndLoop
%	\State Compute $\alpha_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1_{i,j}}W_{1_{i,j}}^2$,
%	$i=1,\ldots,d$.
%    \EndProcedure
%  \end{algorithmic}
%  \label{alg:free1}
%\end{breakablealgorithm}
%\bigskip
%
\bigskip
\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{Algorithm for constructing the matrix, $\mat{C}$}
  \begin{algorithmic}[1]
	\Procedure{Local Linear Approximation}{} 
	\State Draw $N$ random samples, $\{\bm{\xi}_j\}_{j=1}^{N}$ 
	according to $\pi_{\bm{\xi}}$.
	\iffalse \State Project to the physical space:
	$\{\bm{\theta}_k\}_{k=1}^{N}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{N}$ \fi
	\State Compute $f(\bm\xi_j)$, $j=1, \ldots, N$.
	\State Draw $M$ random samples, $\{\bm{\zeta}_i\}_{i=1}^{M}$
	according to $\pi_{\bm{\xi}}$.
	\State Choose an integer $p \leq N$ 
	\State For each $i=1, \ldots, M$, compute 
	\[
	\begin{aligned}
	\Phi_i &= \{ p \text{ nearest points in } \{\bm{\xi}_j\}_{j=1}^{N} \text{ to } \bm{\zeta}_i\}\\
	\vspace{-2mm}
	\Psi_i &= \text{subset of } f(\bm\xi_j) \text{ corresponding to the points in } \Phi_i
	\end{aligned}
	\]
	\State Compute linear least square fits 
	 $f(\bm\xi_j) \approx c_i + b_i^T\bm{\xi}_j$, \  $\bm{\xi}_j \in \Phi_i, \ f(\bm\xi_j) \in \Psi_i$
	\State Compute the matrix 
	\[
	\hat{\mat{C}} \approx \frac{1}{M} \sum_{i=1}^{M} b_i b_i^T = \hat{\mat{W}}\hat{\mat{\Lambda}}\hat{\mat{W}}^\top
	\]
	\EndProcedure
  \end{algorithmic}
  \label{alg:free2}
\end{breakablealgorithm}
\bigskip

>>>>

In this section, we present a generalized framework for computing the
active subspace. The framework is employed to the H$_2$/O$_2$ reaction kinetics
problem whereby the 
pre-exponent ($A_i$) in the rate law associated with individual reactions provided in
Table~\ref{tab:kinetics} is considered to be uniformly
distributed in the interval, $[0.9A_i^\ast, 1.1A_i^\ast]$; $A_i^\ast$ is the nominal
estimate provided in~\cite{Yetter:1991}.
Specifically, we employ two strategies --
a gradient-based strategy that involves computation of model
gradients in order to construct the matrix $\hat{\mat{C}}$
in~\eqref{eq:chat}, and a gradient-free approach 
that involves a regression-based local linear approximation of
the model output to circumvent the computational cost associated with gradient
computation. 
The gradient-based approach is based on Algorithm
1.1, and the gradient-free approach is based on Algorithm 1.2
in~\cite{Constantine:2015}. The two strategies are implemented in an
iterative manner to ensure computational efficiency by avoiding 
unnecessary model evaluations once a converged active subspace
is achieved.

@@

\bigskip
\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{An iterative gradient-based approach for discovering the active subspace}
  \begin{algorithmic}[1]
\Require $\theta_l$, $\theta_u$, $\tau$. 
\Ensure $\hat{\mat{\Lambda}}$, $\hat{\mat{W}}$, $\bm{\nu}_r(f)$ %$\eta$. %
    \Procedure{Gradient-based}{}
    \State Set $k$ = 0
	\State Draw $n_k$ random samples, $\{\bm{\xi}_i\}_{i=1}^{n_k}$ 
         according to $\pi_{\bm{\xi}}$. 
    \State Set $N_\text{total}$ = $n_k$ 
	\iffalse \State Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_r}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_r}$ \fi
	\State For each $i=1, \ldots, N_\text{total}$, compute $f(\bm{\xi}_i)$ and the gradient $\bm{g}^i = \nabla_{\bm{\xi}}f(\bm{\xi}_i)$
    \iffalse  
	\Statex\hspace{5mm} Using Finite Difference: 
	\Statex\hspace{5mm} i. Assign a small increment, $d\xi$.
	\Statex\hspace{5mm} ii. Augment the set of samples with neighboring points.
	\be \{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}:~\{\bm{\xi}_k\}_{k=1}^{n_1} \cup
        \{\xi_{k,j}+d\xi\}_{j=1}^{N_p} \nonumber
	\ee
	\Statex\hspace{5mm} iii. Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}$
	\Statex\hspace{5mm} iv. Using the augmented set, $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}$
         compute $\bm{g}^i$.
         \fi 
	\State Compute $\hat{\mat{C}}$ and its eigenvalue decomposition 
		$\hat{\mat{C}}$= $\frac{1}{N_\text{total}}\sum\limits_{i=1}^{N_\text{total}}[\bm{g}^i][\bm{g}^i]^\top$ = 
		$\hat{\mat{W}}^{(k)}\hat{\mat{\Lambda}}^{(k)} \hat{\mat{W}}^{(k)\top}$
	%\State Eigenvalue decomposition, $\mat{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$%
	\State Partition: $\hat{\mat{\Lambda}}^{(k)}=
        \begin{bmatrix} \hat{\mat{\Lambda}}_1^{(k)} & \\ & \hat{\mat{\Lambda}}_2^{(k)} \end{bmatrix}$, 
        $\hat{\mat{W}}^{(k)}=\begin{bmatrix} \hat{\mat{W}}_1^{(k)} & \hat{\mat{W}}_2^{(k)} \end{bmatrix}$, 
        $\hat{\mat{\Lambda}}_1^{(k)}\in \mathbb{R}^{N_p\times r}$
	\Loop
		\State Set $k$ = $k$ + 1
		\State Draw $n_k =  \lceil\beta n_{k-1}\rceil$  new random samples 
                $\{\bm{\xi}_i\}_{i=1}^{n_k}$  $\beta\in[0,1]$

	%	\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.%
		\State Set $N_\text{total}$ = $N_\text{total}$ + $n_k$ 
		\State Compute $\bm{g}^i = \nabla_{\bm{\xi}_i}f(\bm{\xi}_i)$, 
             	$i=n_{k-1}+1, \ldots, n_{k-1}+n_k$.  
		\State Compute $\hat{\mat{C}}$ = 
        	$\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^i][\bm{g}^i]^\top$
		\State Eigenvalue decomposition, $\hat{\mat{C}}$ = $\hat{\mat{W}}^{(k)}\hat{\mat{\Lambda}}^{(k)}
		 \hat{\mat{W}}^{(k)\top}$
		\State Partition the eigenspace of $\hat{\mat{C}}$ as shown in Step 7
		\State Compute $\delta \hat{\mat{W}}_{1,j}^{(k)}$ = 
                       \scalebox{1.25}{$\frac{\|(\hat{\mat{W}}_{1,j}^{k})^2 - 
                       (\hat{\mat{W}}_{1,j}^{k-1})^2\|}{\|(\hat{\mat{W}}_{1,j}^{k-1})^2\|}$}, 
                       $j = 1,\ldots,r$.
		\If {$\max\left(\delta \hat{\mat{W}}_{1,j}^{(k)}\right)<\tau$}
			\State break
		\EndIf
	\EndLoop
	\State Compute $\nu_{i,r}(f) = \sum\limits_{j=1}^{r} \Lambda_j w_{i,j}^2$,
	$i=1,\ldots,N_p$.
	\State Normalize $\nu_{i,r}(f)$ as $\tilde{\nu}_{i,r}(f)$ = \scalebox{1.25}{$\frac{\nu_{i,r}(f)}{\sum_i\nu_{i,r}(f)}$}.

@@

In the gradient-free approach, the model output is first computed at an initial set of $n_0$ samples. 
An independent set of $M$ samples is also drawn according to the joint probability distribution of the inputs.
For each sample in $M$, $p$ nearest neighbors in the $n_0$ set are identified. A least-squares fit to model
evaluations at the $p$ samples is performed. The slope vector ($\bm{b}_i$) of the regression fit is stored
separately and the process is repeated for each sample in $M$. The symmetric and positive semi-definite matrix,
$\hat{\mat{C}}$ is constructed using the slope vectors as follows:
%
\be
\hat{\mat{C}} = \frac{1}{M}\sum\limits_{i=1}^{M}\bm{b}_i\bm{b}_i^\top = \hat{\bm{W}}\hat{\bm{\Lambda}}\hat{\bm{W}}^\top
\ee
%
Hence, the gradient-free approach essentially differs in the procedure for constructing the matrix, $\hat{\mat{C}}$.
Instead of the model evaluations, a linear regression-based approximation is used to estimate the gradient of
the model output. The suitability of this approach thus depends upon the accuracy of this approximation. 
The active subspace computation involves the same sequence of steps as outlined previously for the gradient-based 
approach in Algorithm~\ref{alg:grad}. The procedure for constructing $\hat{\mat{C}}$ using a regression-based
local linear approximation
of the model output is provided in Algorithm~\ref{alg:free2}.

%\bigskip
%\begin{breakablealgorithm}
%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
%\renewcommand{\algorithmicensure}{\textbf{Output:}}
%  \caption{An iterative gradient-free approach for discovering the active subspace}
%  \begin{algorithmic}[1]
%\Require $\theta_l$, $\theta_u$, $M$, $\tau$. 
%\Ensure $\Lambda$, $W$, $\alpha$ %$\eta$. 
%    \Procedure{Gradient-free}{}
%    \State Set $k$ = 0
%	\State Draw $n_k$ random samples, $\{\bm{\xi}_j\}_{j=1}^{n_k}$
%	\State Set $N_\text{total}$ = $n_k$ 
%	\iffalse \State Project to the physical space:
%        $\{\bm{\theta}_k\}_{i=1}^{n_k}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_r}$
%        \fi
%     \State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ 
%	\State Compute $f(\bm\xi_j)$, $j=1, \ldots, N_\text{total}$.
%	\State Compute the matrix $\mathbb{C}$ and its eigenvalue decomposition using Algorithm 3
%	%\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
%	\State Partition: $\Lambda^{(k)}~=~ 
%        \begin{bmatrix} \Lambda_1^{(k)} & \\ & \Lambda_2^{(k)} \end{bmatrix}$, 
%        $W^{(k)}~=~\begin{bmatrix} W_1^{(k)} & W_2^{(k)} \end{bmatrix}$, 
%        $\Lambda_1^{(k)}\in \mathbb{R}^{d\times\mathcal{S}}$
%	\Loop
%		\State Set $k$ = $k$ + 1
%		\State Draw $n_k =  \beta n_{k-1}$  new random samples 
%		$\{\bm{\xi}_j\}_{j=1}^{n_k}$  $\beta =??$  $j = n_{k-1}+1,\ldots,n_{k-1}+n_k$. 
%        \State $N_\text{total}$ = $N_\text{total}$ + $n_k$    
%	\iffalse	\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$. \fi
%		 
%		\State Compute $f(\bm\xi_j)$, $j=n_{k-1}+1, \ldots, n_{k-1}+n_k$.  
%		\State Construct the matrix, $\mat{C}$ and its eigenvalue decomposition
%		\State Compute $\delta W_{1,j}^{(k)}$ = 
%                       \scalebox{1.25}{$\frac{\|(W_{1,j}^{k})^2 - 
%                       (W_{1,j}^{k-1})^2\|}{\|(W_{1,j}^{k-1})^2\|}$}, 
%                       $j = 1,\ldots,\mathcal{S}$.
%		\If {$\max\left(\delta W_{1,j}^{(k)}\right)<\tau$}
%			\State break
%		\EndIf
%	\EndLoop
%	\State Compute $\alpha_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1_{i,j}}W_{1_{i,j}}^2$,
%	$i=1,\ldots,d$.
%    \EndProcedure
%  \end{algorithmic}
%  \label{alg:free1}
%\end{breakablealgorithm}
%\bigskip
%
\bigskip
\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{Algorithm for constructing the matrix, $\hat{\mat{C}} in~\eqref{eq:chat}$}
  \begin{algorithmic}[1]
	\Procedure{Local Linear Approximation}{} 
	\State Draw $N$ random samples, $\{\bm{\xi}_j\}_{j=1}^{N}$ 
	according to $\pi_{\bm{\xi}}$.
	\iffalse \State Project to the physical space:
	$\{\bm{\theta}_k\}_{k=1}^{N}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{N}$ \fi
	\State Compute $f(\bm\xi_j)$, $j=1, \ldots, N$.
	\State Draw $M$ random samples, $\{\bm{\zeta}_i\}_{i=1}^{M}$
	according to $\pi_{\bm{\xi}}$.
	\State Choose an integer $p \leq N$ 
	\State For each $i=1, \ldots, M$, compute 
	\[
	\begin{aligned}
	\Phi_i &= \{ p \text{ nearest points in } \{\bm{\xi}_j\}_{j=1}^{N} \text{ to } \bm{\zeta}_i\}\\
	\vspace{-2mm}
	\Psi_i &= \text{subset of } f(\bm\xi_j) \text{ corresponding to the points in } \Phi_i
	\end{aligned}
	\]
	\State Compute linear least square fits 
	 $f(\bm\xi_j) \approx c_i + b_i^T\bm{\xi}_j$, \  $\bm{\xi}_j \in \Phi_i, \ f(\bm\xi_j) \in \Psi_i$
	\State Compute the matrix 
	\[
	\hat{\mat{C}} \approx \frac{1}{M} \sum_{i=1}^{M} b_i b_i^T = \hat{\mat{W}}\hat{\mat{\Lambda}}\hat{\mat{W}}^\top
	\]
	\EndProcedure
  \end{algorithmic}
  \label{alg:free2}
\end{breakablealgorithm}
\bigskip
